---
title: "Mariners 2022"
author: "Nicholas Kondo"
subtitle: 
output:
  html_document:
    df_print: paged
    toc: true 
  html_notebook: default
---

```{r setup, include= FALSE}
library(knitr)

# Change the number in set seed to your own favorite number
set.seed(4)
options(width=70)
options(scipen=99)

# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = TRUE, 
               cache.lazy = FALSE,
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')
```

```{r setup-2, include= FALSE}

# Always print this out before your assignment
sessionInfo()
getwd()

```

```{r setup-3, include =FALSE}
# Loading libraries and data 

library(here)
library(dplyr)
library(ggplot2)
library(forcats)
library(viridis)
library(randomForest)
library(mgcv)
library(rpart)
library(partykit)
library(pROC)
library(caret)
library(tidymodels)
library(DMwR)
library(ROSE)

train <- read.csv(here('Datasets','train.csv'))

```

# 1.  Data Exploration 

```{r, include = FALSE}
# Exploring variable names
names(train)
```

The variable we are trying to predict, home runs, occurs less than 1% of all pitches.  
```{r}
table(train$is_hr)

homeruns <- train %>% 
  filter(is_hr == 1)
```


### 1.1 Pitch Location
Let's look at how the location of pitches affects the probability of hitting a home run.  This heat map aligns with our baseball inference that pitches closer to the middle of the zone are more likely to be hit for a home run.  It'd be interesting to look at heat maps for different types of pitches and separately for RHP and LHP.  Plots for these splits are located in the [Appendix](#appendix).
```{r}
# Creating generalized model
      fit <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = train)
      # Finding predicted probabilities over a 50 x 50 grid
      x <- seq(-1.5, 1.5, length.out=50)
      y <- seq(0.5, 5, length.out=50)
      data.predict <- data.frame(plate_side = c(outer(x, y * 0 + 1)),
                                 plate_height = c(outer(x * 0 + 1, y)))
      # Creating LP model
      lp <- predict(fit, data.predict)
      # Adding the probability of hitting a home run given the location 
      data.predict$Probability <- exp(lp) / (1 + exp(lp))
      # Creating the K Zone
      topKzone <- 3.5
      botKzone <- 1.6
      inKzone <- -0.95
      outKzone <- 0.95
      kZone <- data.frame(
        x=c(inKzone, inKzone, outKzone, outKzone, inKzone),
        y=c(botKzone, topKzone, topKzone, botKzone, botKzone))
      # Constructing the plot for ALL pitchers
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= Probability)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "All Pitchers",
             caption = paste("N = ", nrow(train)))
```
 
 ### 1.2 Missing Values
Exploring the missing values across the data set.  
```{r}
sort(colSums(is.na(train)))
```

There are a lot of NA values for pitch type... let's check out the different types of pitches.  
```{r}
train$pitch_type <- as.factor(train$pitch_type)
table(train$pitch_type)
```

There are six types of pitches listed in this data set with a majority of pitches being  a fastball, and a low number of knuckleballs.  Here we are going to look at a dataset of the pitches where `pitch_type = NA`
```{r}
NA_pitchType <- train %>% 
  filter(is.na(train$pitch_type))
```
 
 
 
# 2. Feature Engineering

## 2.1 Handing Missing Values 

### Pitch Type 
After scrolling through the data set containing NA values, which can be looked at here **hyperlink**, the pitches seem to be somewhat normal- the pitches weren't spike or thrown over the backstop, and some actually were even hit for home run.  We should try to keep this data.  My next thought is that only so many pitches can fall under the six classes of pitches in the data set.  

I would assume two-seamers and cut-fastballs could fall under fastballs, but what about a pitch like a knuckle curve?  Would that be considered a curve or a knuckball?  Perhaps there are pitches that were not able to be classified.  This data should be preserved so for all pitches listed NA, it will be classified as `other`

**Provide link to dataset**

```{r}
# Changing pitch type to character, replacing NA values with "other", then converting back to factor
train1 <- train  %>% 
  select(-date, -umpire_id, -catcher_id,  -inning, -top_bottom, -y55, -pitch_id, -pitcher_id, -batter_id, -tilt)

train1$pitch_type <- as.character(train1$pitch_type)
train1$pitch_type[is.na(train1$pitch_type)] <- "Other"
train1$pitch_type <- as.factor(train1$pitch_type)

# Confirming NA values
sort(colSums(is.na(train1)))
levels(train1$pitch_type)

```

### Spin Rate
To accomodate for spin rate missing values, we are going to impute the median spin rate based on the type of pitch.  To do this first we need to separate the pitches by pitch type, compute the median, then impute the median for the missing values.  
```{r}
# Creating separate data sets for each pitch type 
fastballs <- train1 %>% 
  filter(pitch_type == 'FA')
curveballs <- train1 %>% 
  filter(pitch_type == 'CU')
changeups <- train1 %>% 
  filter(pitch_type == 'CH')
sinkers <- train1 %>% 
  filter(pitch_type == 'SI')
sliders <- train1 %>% 
  filter(pitch_type == 'SL')
knuckleballs <- train1 %>% 
  filter(pitch_type == 'KN')
others <- train1 %>% 
  filter(pitch_type == 'Other')

# Filling all NA spin_rate values with the median spin rate of that respective pitch
fastballs$spin_rate[is.na(fastballs$spin_rate)] <- median(fastballs$spin_rate, na.rm=TRUE)
curveballs$spin_rate[is.na(curveballs$spin_rate)] <- median(curveballs$spin_rate, na.rm=TRUE)
changeups$spin_rate[is.na(changeups$spin_rate)] <- median(changeups$spin_rate, na.rm=TRUE)
sinkers$spin_rate[is.na(sinkers$spin_rate)] <- median(sinkers$spin_rate, na.rm=TRUE)
sliders$spin_rate[is.na(sliders$spin_rate)] <- median(sliders$spin_rate, na.rm=TRUE)
knuckleballs$spin_rate[is.na(knuckleballs$spin_rate)] <- median(knuckleballs$spin_rate, na.rm=TRUE)
others$spin_rate[is.na(others$spin_rate)] <- median(others$spin_rate, na.rm=TRUE)

# Combining our pitch types back together
train1 <- bind_rows(fastballs, curveballs, changeups, sinkers, sliders, knuckleballs, others)

rm(fastballs, curveballs, changeups, sinkers, sliders, knuckleballs, others) # Freeing memory

sort(colSums(is.na(train1)))
```

### Other Missing/Usual Values
At this stage, we don't have many more missing values.  There are a series of unusual values where `vert_break`, `induced_break`, and `horz_break` have values of 0.0000.  We will remove the remaining unsual and missing values. 

```{r}
train1 %>% 
  filter(is.na(vert_break))
train1 %>% 
  filter(vert_break == 0)
# I find it very wierd that we are seeing values of 0.000 and I cannot explain the NA values for break, it may be random
# Break is a very important variable so to access this accurately we should just throw out these missing values 
# There is about 171 rows with NA and 450 rows of 0's across the board
# I am ok with throwing away the data 
# PCA?

# Dropping NA values
train1 <- na.omit(train1)

# dropping rows where break was not recorded or recorded as 0.0000
train1 <-subset(train1, vert_break!= 0)

train1 <- as.data.frame(unclass(train1),                     
                        stringsAsFactors = TRUE)
```

## 2.2 Resampling for the Imbalanced Classes 

Since a home run occurs less than 1% of all pitches in the data set, a model like a random forest can predict ‘No HR’ every observation and the model would be correct 99% of the time, but that’s not useful for us.  Mulitple techniques for resolving the class imbalances were considered such as down-sampling, up-sampling, and hybrid method algorithms like SMOTE and ROSE.  This is different from splitting the data set.  The data was split into training and testing.  Once a balanced data set was created from the unbalanced training set, a prediction model was built off this augmented data set, and applied to the unbalanced test data.  

### Down-sampling
The first resampling technique was down-sampling.  Down sampling is where we randomly subset all the classes in the training set so that their class frequencies match the minority class.  This would result in using only 2% of our training data being used to fit the mode.  
```{r}
# Let's create different versions of the training set prior to model tuning

# Down sampling the data set 

    # Changing is_hr, balls, strikes, and outs to a factor
    train2 <- train1 %>%
      mutate(is_hr = ifelse(is_hr == 1, "Class2", "Class1")) %>% 
      mutate_if(is.character, factor) 

    train2 <- train2 %>% 
      rename(Class = is_hr)

pitches_split <- initial_split(train2, strata = Class)
pitches_train <- training(pitches_split)
pitches_test <- testing(pitches_split)

# Clearing up memory 
rm(train, train1)
```


```{r}
# Down sampling the data set 
set.seed(400)
down_train <- downSample(x = pitches_train[, -ncol(pitches_train)],
                         y = pitches_train$Class)

table(down_train$Class)
```

### Up-sampling
The next resampling method is up-sampling.  Up-sampling randomly samples the rare-occurring class to be the same size as the majority class.  In other words, we will now have the same amount of home runs as non-home runs.  
```{r}
# Up sampling the data
set.seed(400)
up_train <- upSample(x = pitches_train[, -ncol(pitches_train)],
                     y = pitches_train$Class)                         
table(up_train$Class) 
```

### SMOTE
SMOTE stands for *synthetic minority over-sampling technique.  This technique synthesizes new occurrences of home runs from the existing examples. 
```{r}
# SMOTE
set.seed(9560)
smote_train <- SMOTE(Class ~ ., data  = pitches_train)                         
table(smote_train$Class) 
```

### ROSE
ROSE stands for random over-sampling examples and is a bootstrap-based technique to help deal with the imbalanced classes.  
```{r}
# ROSE
set.seed(9560)
rose_train <- ROSE(Class ~ ., data  = pitches_train)$data                         
table(rose_train$Class) 
```

### Cross Validation
To evaluate the techniques  we will used a bagged classification and estimate the AUC (Area under ROC Curve) using five repeats of 10 fold cross validation.  
```{r}
ctrl <- trainControl(method = "repeatedcv", repeats = 5,
                     classProbs = TRUE,
                     summaryFunction = twoClassSummary)

set.seed(5627)
orig_fit <- train(Class ~ ., data = pitches_train, 
                  method = "treebag",
                  nbagg = 10,
                  metric = "ROC",
                  trControl = ctrl)
```
```{r}
set.seed(562)
down_outside <- train(Class ~ ., data = down_train,
                      method = "treebag",
                      nbagg = 10,
                      metric = "ROC",
                      trControl = ctrl)
```
```{r}
set.seed(562)
up_outside <- train(Class ~ ., data = up_train, 
                    method = "treebag",
                    nbagg = 10,
                    metric = "ROC",
                    trControl = ctrl)
```
```{r}
set.seed(562)
rose_outside <- train(Class ~ ., data = rose_train, 
                      method = "treebag",
                      nbagg = 10,
                      metric = "ROC",
                      trControl = ctrl)
```

## Appendix {#appendix}

###Heat maps

[RHP - All Pitch Types](#RHPAllpitches)

[LHP - All Pitch Types](#LHPAllpitches)

[RHP - Fastballs](#RHPFA)

[LHP - Fastballs](#LHPFA)

[RHP - Curveballs](#RHPCU)

[LHP - Curveballs](#LHPCU)

[RHP - Sliders](#RHPSL)

[LHP - Sliders](#LHPSL)

[RHP - Changeups](#RHPCH)

[LHP - Changeups](#LHPCH)

[RHP - Sinkers](#RHPSI)

[LHP - Sinkers](#LHPSI)

[RHP - Knuckleballs](#RHPKN)

[LHP - Knuckleballs](#LHPKN)

Home run probabilities for all pitch types from Right-Handed Pitchers. {#RHPAllpitches}
```{r, echo = FALSE}
 # Creating data set with only RHP
      RHP <- train %>% 
        filter(pitcher_side == 'Right')
      # Creating data set with only LHP
      LHP <- train %>% 
        filter(pitcher_side == 'Left')
      # Creating generalized model
      fitRHP <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = RHP)
      fitLHP <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = LHP)
      # Finding predicted probabilities over a 50 x 50 grid
      x <- seq(-1.5, 1.5, length.out=50)
      y <- seq(0.5, 5, length.out=50)
      data.predict <- data.frame(plate_side = c(outer(x, y * 0 + 1)),
                                 plate_height = c(outer(x * 0 + 1, y)))
      # Creating LP model
      lpRHP <- predict(fitRHP, data.predict)
      lpLHP <- predict(fitLHP, data.predict)
      # Adding the probability of hitting a home run given the location 
      data.predict$ProbabilityRHP <- exp(lpRHP) / (1 + exp(lpRHP))
      data.predict$ProbabilityLHP <- exp(lpLHP) / (1 + exp(lpLHP))
      
      # Constructing plot
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityRHP)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "RHP - All Pitches",
             caption = paste("N = ",nrow(RHP)))
```

Home run probabilities for all pitch types from Left-Handed Pitchers. {#LHPAllpitches}
```{r, echo = FALSE}
 ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityLHP)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "LHP - All Pitches",
             caption = paste("N = ",nrow(LHP)))
```

```{r, include = FALSE}
 -------------------------------------------------------------------------------
        # Creating heat map for each pitch type (RHP)
      -------------------------------------------------------------------------------
      ## Creating Data sets 
      # Fastballs
      RHP_FA <- train %>% 
        filter(pitcher_side == 'Right',
               pitch_type == "FA")
      # Curveballs
      RHP_CU <- train %>% 
        filter(pitcher_side == 'Right',
               pitch_type == "CU")
      # Sliders
      RHP_SL <- train %>% 
        filter(pitcher_side == 'Right',
               pitch_type == "SL")
      # Changeups
      RHP_CH <- train %>% 
        filter(pitcher_side == 'Right',
               pitch_type == "CH")
      # Sinkers
      RHP_SI <- train %>% 
        filter(pitcher_side == 'Right',
               pitch_type == "SI")
      # Knuckleballs
      RHP_KN <- train %>% 
        filter(pitcher_side == 'Right',
               pitch_type == "KN")
      
       -------------------------------------------------------------------------------
        # Creating heat map for each pitch type (LHP)
      -------------------------------------------------------------------------------
      
      ## Creating Data sets 
      # Fastballs
      LHP_FA <- train %>% 
        filter(pitcher_side == 'Left',
               pitch_type == "FA")
      # Curveballs
      LHP_CU <- train %>% 
        filter(pitcher_side == 'Left',
               pitch_type == "CU")
      # Sliders
      LHP_SL <- train %>% 
        filter(pitcher_side == 'Left',
               pitch_type == "SL")
      # Changeups
      LHP_CH <- train %>% 
        filter(pitcher_side == 'Left',
               pitch_type == "CH")
      # Sinkers
      LHP_SI <- train %>% 
        filter(pitcher_side == 'Left',
               pitch_type == "SI")
      # Knuckleballs
      LHP_KN <- train %>% 
        filter(pitcher_side == 'Left',
               pitch_type == "KN")
```

```{r, include = FALSE}
# Creating generalized models (RHP)
      fitRHP_FA <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = RHP_FA)
      fitRHP_CU <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = RHP_CU)
      fitRHP_SL <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = RHP_SL)
      fitRHP_CH <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = RHP_CH)
      fitRHP_SI <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = RHP_SI)
      fitRHP_KN <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = RHP_KN)
      
# Creating generalized models (LHP)
      fitLHP_FA <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = LHP_FA)
      fitLHP_CU <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = LHP_CU)
      fitLHP_SL <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = LHP_SL)
      fitLHP_CH <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = LHP_CH)
      fitLHP_SI <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = LHP_SI)
      fitLHP_KN <- gam(is_hr ~ s(plate_side, plate_height), family = binomial, data = LHP_KN)
```

```{r, INCLUDE = FALSE}
# Finding predicted probabilities over a 50 x 50 grid
      x <- seq(-1.5, 1.5, length.out=50)
      y <- seq(0.5, 5, length.out=50)
      data.predict <- data.frame(plate_side = c(outer(x, y * 0 + 1)),
                                 plate_height = c(outer(x * 0 + 1, y)))
      
      # Creating LP models
      lpRHP_FA <- predict(fitRHP_FA, data.predict)
      lpRHP_CU <- predict(fitRHP_CU, data.predict)
      lpRHP_SL <- predict(fitRHP_SL, data.predict)
      lpRHP_CH <- predict(fitRHP_CH, data.predict)
      lpRHP_SI <- predict(fitRHP_SI, data.predict)
      lpRHP_KN <- predict(fitRHP_KN, data.predict)
      
       # Creating LP models
      lpLHP_FA <- predict(fitLHP_FA, data.predict)
      lpLHP_CU <- predict(fitLHP_CU, data.predict)
      lpLHP_SL <- predict(fitLHP_SL, data.predict)
      lpLHP_CH <- predict(fitLHP_CH, data.predict)
      lpLHP_SI <- predict(fitLHP_SI, data.predict)
      lpLHP_KN <- predict(fitLHP_KN, data.predict)

```

```{r, include = FALSE}
 # Adding the probability of hitting a home run given the location (RHP)
      data.predict$ProbabilityRHP_FA <- exp(lpRHP_FA) / (1 + exp(lpRHP_FA))
      data.predict$ProbabilityRHP_CU <- exp(lpRHP_CU) / (1 + exp(lpRHP_CU))
      data.predict$ProbabilityRHP_SL <- exp(lpRHP_SL) / (1 + exp(lpRHP_SL))
      data.predict$ProbabilityRHP_CH <- exp(lpRHP_CH) / (1 + exp(lpRHP_CH))
      data.predict$ProbabilityRHP_SI <- exp(lpRHP_SI) / (1 + exp(lpRHP_SI))
      data.predict$ProbabilityRHP_KN <- exp(lpRHP_KN) / (1 + exp(lpRHP_KN))
      
       # Adding the probability of hitting a home run given the location (LHP)
      data.predict$ProbabilityLHP_FA <- exp(lpRHP_FA) / (1 + exp(lpLHP_FA))
      data.predict$ProbabilityLHP_CU <- exp(lpRHP_CU) / (1 + exp(lpLHP_CU))
      data.predict$ProbabilityLHP_SL <- exp(lpRHP_SL) / (1 + exp(lpLHP_SL))
      data.predict$ProbabilityLHP_CH <- exp(lpRHP_CH) / (1 + exp(lpLHP_CH))
      data.predict$ProbabilityLHP_SI <- exp(lpRHP_SI) / (1 + exp(lpLHP_SI))
      data.predict$ProbabilityLHP_KN <- exp(lpRHP_KN) / (1 + exp(lpLHP_KN))
```

Home run probabilities on Fastballs from Right-Handed Pitchers{#RHPFA}
```{r, echo = FALSE}
 # Constructing GGplots
      # Fastballs LHP
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityRHP_FA)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "RHP Fastballs",
             caption = paste("N = ",nrow(RHP_FA)))
```

Home run probabilities on Fastballs from Left-Handed Pitchers{#LHPFA}
```{r, echo = FALSE}
 # Constructing GGplots
      # Fastballs LHP
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityLHP_FA)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "LHP Fastballs",
             caption = paste("N = ",nrow(LHP_FA)))
```

Home run probabilities on Curveballs from Right-Handed Pitchers{#RHPCU}
```{r, echo = FALSE}
# Curveballs
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityRHP_CU)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "RHP Curveballs",
             caption = paste("N = ",nrow(RHP_CU)))
```

Home run probabilities on Curveballs from Left-Handed Pitchers{#LHPCU}
```{r, echo = FALSE}
# Curveballs
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityLHP_CU)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "LHP Curveballs",
             caption = paste("N = ",nrow(LHP_CU)))
```

Home run probabilities on Sliders from Right-Handed Pitchers{#RHPSL}
```{r, echo = FALSE}
# Sliders
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityRHP_SL)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "RHP Sliders",
             caption = paste("N = ",nrow(RHP_SL)))
```

Home run probabilities on Sliders from Left-Handed Pitchers{#LHPSL}
```{r, echo = FALSE}
# Sliders
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityLHP_SL)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "LHP Sliders",
             caption = paste("N = ",nrow(LHP_SL)))
```

Home run probabilities on Changeups from Right-Handed Pitchers{#RHPCH}
```{r, echo = FALSE}
# Change Ups
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityRHP_CH)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "RHP Change Ups",
             caption = paste("N = ",nrow(RHP_CH)))
```

Home run probabilities on Changeups from Left-Handed Pitchers{#LHPCH}
```{r, echo = FALSE}
# Change Ups
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityLHP_CH)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "LHP Change Ups",
             caption = paste("N = ",nrow(LHP_CH)))
```

Home run probabilities on Sinkers from Right-Handed Pitchers{#RHPSI}
```{r, echo = FALSE}
# Sinkers
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityRHP_SI)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "RHP Sinkers",
             caption = paste("N = ",nrow(RHP_SI)))

```

Home run probabilities on Sinkers from Left-Handed Pitchers{#LHPSI}
```{r, echo = FALSE}
# Sinkers
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityLHP_SI)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "LHP Sinkers",
             caption = paste("N = ",nrow(LHP_SI)))
```

Home run probabilities on Knuckleballs from Right-Handed Pitchers{#RHPKN}.  Keep in mind the sample size is small.  
```{r, echo = FALSE}
 # Knuckleballs
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityRHP_KN)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "RHP Knuckleballs",
             caption = paste("N = ",nrow(RHP_KN)))
```

Home run probabilities on Knuckleballs from Left-Handed Pitchers{#LHPKN}.  Keep in mind the sample size is small.  
```{r, echo = FALSE}
 # Knuckleballs
      ggplot(kZone, aes(x, y)) +
        geom_tile(data=data.predict, 
                  aes(x= plate_side, y= plate_height, fill= ProbabilityLHP_KN)) +
        scale_fill_distiller(palette = "Spectral") +
        geom_path(lwd=1.5, col="black") +
        coord_fixed() + 
        labs(title = "HR Probability Based on Pitch Location",
             subtitle = "LHP Knuckleballs",
             caption = paste("N = ",nrow(LHP_KN)))
```

```{r, echo = FALSE}


```

```{r, echo = FALSE}


```

```{r, echo = FALSE}


```

```{r, echo = FALSE}


```

```{r, echo = FALSE}


```



`date`: Date in MM/DD/YY

`pitcher_id`: Pitcher identifier, unique

`pitcher_side` Handedness of pitcher (right or left)

`batter_id`: Batter identifier, unique

`batter_side`: Handedness of batter (right or left)

`stadium_id`: Stadium identifier, unique

`umpire_id`: Umpire identifier, unique

`catcher_id`: Catcher identifier, unique

`inning`: Inning (1-9)

`top_bottom`: Top or Bottom of the inning. (1 or 2, 1 = ? 2 + ?)

`outs`: Number of outs (0, 1, or 2)

`balls` Number of balls (0, 1, 2, or 3)

`strikes` Number of strikes (0, 1, or 2)

`release_speed`: Speed of pitch when it leaves the pitcher’s hand

`vert_release_angle`: Initial vertical (up-down) direction of the ball when it leaves the pitcher’s hand, reported in degrees. A positive number means the ball is released upward, while a negative number means the ball is released downward

`horz_release_angle`: Initial horizontal (left-right) direction of the ball when it leaves the pitcher’s hand, reported in degrees. A positive number means the ball is released to the right from the pitcher’s perspective, while a negative number means the ball is released to the left from the pitcher’s perspective.

`spin_rate`: How fast the ball is spinning aas it leaves the pitcher's hand, reported in the number of times the pitched ball would spin per minute (revolutions per minute)

`spin_axis`: Direction the ball is spinning, reported in degrees of tilt. Note that: 
A ball with a spin axis of 0 has pure top spin.
A ball with a spin axis of 180 has pure backspin.  The ball drops less than what gravity would cause alone.
A ball with a spin axis of 90 is spinning squarely toward the left from a pitcher's perspective and will break to the left.
A ball with a spin axis of 270 is spinning squarely toward the right, from a pitcher's perspective and will break to the right.

`tilt`: Spin axis converted into clock time, rounded to the nearest 15 minutes. As a rule of thumb, the ball will break in the direction of the number on the clock face. For example:
a. 6:00 is perfect top spin (classic “12 – 6” curveball), causing the ball to break down
b. 12:00 is perfect back spin (Four seam fastball, with no left-right movement), causing the ball to break upward relative to how it would have moved due to gravity alone - cutters are around 11:00 and sinkers are around 2:00 for a RHP, while cutters are around 1:00 and sinkers around 10:00 for a LHP
c. 3:00 is a “Frisbee” spinning and breaking to the right, while 9:00 is a “Frisbee” spinning and breaking to the left.

`rel_height`: Height of the ball 

`rel_side`

`extension`: Distance from the rubber at which the pitcher releases the ball

`vert_break`: Distance between where the pitch actually crosses the front of home plate height-wise, and where it would have crossed home plate height-wise if had it traveled in a perfectly straight line from release, completely unaffected by gravity. Note: This number will be quite large for pitches released with a positive vertical angle

`induced_vert_break`: Distance between where the pitch actually crosses the front of home plate height-wise, and where it would have crossed home plate height-wise if had it traveled in a perfectly straight line from release, but affected by gravity. Note: If this number is positive, the ball broke “upwards”, or in reality dropped less than it would have due to gravity alone – it does not necessarily mean that the ball actually rose.

`horz_break`: Distance between where the pitch actually crosses the front of home plate sidewise, and where it would have crossed home plate side-wise if had it traveled in a perfectly straight line from release. A positive number means the break was to the right from the pitcher’s perspective, while a negative number means the break was to the left from the pitcher’s perspective.

`plate_height`

`plate_side`

`zone_speed`: Speed of pitch when it crosses the plate

`vert_approach_angle`:

`horz_approach_angle`:

`x55`

`y55`

`z55`

`pitch_type`:  Pitch classification (Fastball, Curveball, etc.)

`pitch_id`: Pitch identifier

`is_hr`: Binary indicator is pitch is home run or not.  

references: https://emspeedtraining.com/programs/driveline/
